---
title: "Capital"
author:
  - Vasileios Papadopoulos
  
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages('rpart')
# library('rpart')
```

## Dataset
Dataset consists of two features. The numeric variable *balance* and binary variable *gender*. 

```{r}
capital <- read.table('capital.csv', header = TRUE, sep = ';')
attach(capital)
head(capital)
head(is.na.data.frame(capital))
```

## Relative frequency
We will explore *balance* in relation to *gender*. Particularly, we are interested in frequency tables, bar and pie plots. Frequency table is defined as:

\begin{equation} 
  \hat{\text{Relative Frequency}} = \frac{\text{Frequency}}{\text{Sample Size}}
\end{equation}

Get possible values of *gender*

```{r}
factor(capital$gender)
```

```{r}
gender.freq = table(capital$gender)
gender.relfreq = gender.freq / nrow(capital)
```

```{r relative-freq-plots, fig.show="hold", out.width="50%"}
plot(gender.relfreq)
pie(gender.relfreq)
```

```{r box-plots, fig.show="hold", out.width="50%"}
balance = capital$balance
balance_1 = capital[capital$gender == 1]
balance_2 = capital[capital$gender == 2]
boxplot(capital$balance)
boxplot(balance_1)
boxplot(balance_2)
```



