---
title: "Fuel consumption difference between Automatic and Manual transmission"
author:
  - Georgios Papadopoulos
  - Vasileios Papadopoulos
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
head(mtcars)
```

Check whether *mpg* comes from a Gaussian distribution with Q-Q plot.

```{r mpg, fig.show="hold", out.width="50%"}
hist(mtcars$mpg)
qqnorm(mtcars$mpg, pch = 1)
qqline(mtcars$mpg, lwd = 2)
```

Binary feature *am* represents *automatic* with 0 and *manual* with 1. For convenience we convert it into a categorical variable.
```{r} 
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <- c('at','mt')
head(mtcars)
```

We split the dataset into two subsets *Automatic(AT)* and *Manual (MT)* transmission. We want to find confidence intervals for 0.95 confidence for the mean *MPG* for both categories. In order to do this we run *t-test*

```{r}
mpg.automatic <- mtcars[mtcars$am == "at",]$mpg
mpg.manual <- mtcars[mtcars$am == "mt",]$mpg
t.test(mpg.automatic, mpg.manual)
```

P-value is *0.001374* which is below the 5% confidence, so alternative hypothesis is accepted. The 95% confidence interval of the difference in mean fuel consumption between *Automatic* and *Manual* transmission is between *3.2* and *11.2*






