---
title: "Difference between hair color in Females and Males"
author:
  - Georgios Papadopoulos
  - Vasileios Papadopoulos
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

Given the following dataset test whether there is a statistical difference between males and females on hair color.

```{r}
hair_color <- read.table('hair_color.txt', header = TRUE, sep = ',', row.names = 1)
attach(hair_color)
head(hair_color)
head(is.na.data.frame(hair_color))
```



## Hypothesis

We have two independent populations grouped by gender. We would like to check whether gender plays any role in hair color of individuals. We set the hypothesis as follow:

In Null hypothesis we assume that there is no effect whatsoever
H0 : Gender has no effect on hair color. There is no statistical difference in two gender groups.
Ha : Gender has effect on hair color. There is a statistical difference in two gender groups.

## Statistical tests

Since we have categorial data we use Chi-squared. Alternately, we could have used Fisher's test but the datase is too small.

### Chi-squared test

```{r}
tests.chi <- chisq.test(hair_color)
tests.chi
```

We observe that the *p-value = 0.0325* which is smaller than the significance level *a = 0.05* thus, we fail to refect the null hypothesis.