---
title: "Insurance"
author:
  - Georgios Papadopoulos
  - Vasileios Papadopoulos
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("psych",dependencies=TRUE)
```


## Load Dataset

For the purpose of the exercise we will use the package *psych*. Psych is a package developed for personality, psychometric and psychometric research. It provides useful functions for such analysis and it is a core part of International Cognitive Ability Resource (ICAR) project[1].

Dataset consists of 1338 records and 7 features. The column *charges* is the independent variable, the other 6 will be used to analyze their impact to total costs.

```{r}
require(psych)
df <- read.csv(file = 'insurance.csv')
nrow(df)
ncol(df)
summary(df)
```

## Describe dataset

Now we will use the *describe* method provided by psych package. It let us for a more in depth overview of the data by presenting the frequently used descriptive statistics for psychometric and psychology research. Note the symbol * indicates that the variable is categorical.

```{r}
describe(df)
```

We can see about the *mean*, *standard deviation*, *median*, *trimmed*, *mean absolute deviation*, *min*, *max*, *range*, *skew*, *kurtosis* and *standard error*. Before proceeding to model construction that it could explain/predict the dependent variable(*charges*) we need to define skewness and kyrtosis.

### Skewness

Skewness is described as a measure of data symmetry. A perfectly symmetrical data will have a skewness of 0 which might indicate a Normal distribution as the value of skewness for the latter is also 0. 

Skewness is defined as:

\begin{equation}
a_{3} = \sum{\frac{(X_{i} - \bar{X})^{3}}{ns}}
\end{equation}

where:

- *n* is the sample size
- $X_{i}$ is the $i^{th}$ X value
- $\bar{X}$ is the average
- *s* is the sample standard deviation

The exponent *3* is referred to the third standardized central moment for the probability model.

Usually, we interpret the  its value (rule of thumb) as:

- If the skewness is between -0.5 and 0.5, the data are fairly symmetrical
- If the skewness is between -1 and – 0.5 or between 0.5 and 1, the data are moderately skewed
- If the skewness is less than -1 or greater than 1, the data are highly skewed

### Kyrtosis

Kurtosis is a measure of whether a distribution is narrowly concentrated to the middle; most of the responses are in the center. In other words is a measure of peakedness or flatness of data points.

Kurtosis is defined as:

\begin{equation}
a_{4} = \sum{\frac{(X_{i} - \bar{X})^{4}}{ns}}
\end{equation}

where:

- *n* is the sample size
- $X_{i}$ is the $i^{th}$ X value
- $\bar{X}$ is the average
- *s* is the sample standard deviation

The exponent *4* is referred to the fourth standardized central moment for the probability model.

Analysing the numerical variables of the dataset and the output of *psych.describe* we can see that the variable *bmi* with *skew = 0.28* and *kyrtosis = -0.06* is distributed fairly normally.


```{r hists}
hist(df$age)
hist(df$bmi)
hist(df$children)
hist(df$charges)
```


## Model

We will test two models, *multivariate linear regression* and *decision tree*. For the first case we need transform categorical variables into numerical


### Multivariate Linear Regression
```{r}
str(df)
```

As mentioned previously, there are 3 features which are categorical. In the snipper above we see the *sex*, *smoker* and *region* have a structure of chr. We need to convert them into *Factors* in order to fit a linear regression model.We call the *as.factor* method.

```{r}
df$sex <- as.factor(df$sex)
df$smoker <- as.factor(df$smoker)
df$region <- as.factor(df$region)
str(df)
```

### Fit model

```{r}
linear <- lm(df$charges~df$age + df$bmi + df$children + df$sex + df$smoker + df$region)
summary(linear)
```

## References

psych package [@https://personality-project.org/r/psych/].

