---
title: "Ionoshere"
author:
- Georgios Papadopoulos
- Vasileios Papadopoulos
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
#install.packages("RWeka")
require(RWeka)
```

## Load Dataset

```{r}
df <- read.arff("ionoshere.arff")
head(df)
```

## Methodology
* Define performance metrics
* Cross Validation
* Algorithms(KNN)
* PCA

### Cross Validation

The validation set approach consists of randomly splitting the data into two sets: one set is used to train the model and the remaining other set sis used to test the model. Steps:

1. Traun a model on the training data set
2. Apply the model to the test data set to predict the outcome of new unseen observations
3. Quantify the prediction error, define performance metric

### Performance metric
For our analysis the following performance metrics will be used. 

* Accuracy
* Recall 
* Precision
* F1 - Score

### KNN

We will use k-nearest neighbors to classify the data. KNN is an non-parametric classification algorithm. Since it uses Eucledian distance we will need to scale the attributes.


